CudeleFS: Programmable Consistency and Durability\\in a Global Namespace

### 2017-04-01 paper/sections/4_methodology.tex
paper: integrate noah section 3 feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Methodology: Decoupled Namespaces}
\label{sec:methodology-decoupled-namespaces}

\begin{figure}[tb]
\caption{Applications can decouple the namespace, write updates to a local
journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
these mechanisms (represented by the arrows) can be combined to provide weaker
consistency or durability semantics.  }\label{fig:decouple}
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\end{figure}

In this section we describe Cudele, our prototype system that lets future
programmers compose mechanisms (Section~\S\ref{sec:cudeles-mechanisms}) to
provide the necessary guarantees
(Section~\S\ref{sec:setting-policies-with-cudele}) for their application.

@@

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace. The metadata store represents
the namespace as a tree of directory fragments and is easier to read/traverse.
The journal represent the namespace as a list of events. It is
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more data to
read, it is unorganized, and many of the updates may be redundant.

% describe the mechanisms
Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Local Persist" takes the in-memory journal and writes it to the client's
disk. ``Global Persist" saves the journal as a an object in the object store from the
client.

Next, we discuss how these mechanisms can be composed to get different
consistency and durability semantics.

%\begin{tabular}{ r | l }
%  \(\Rightarrow\)   & Description \\\hline
%  create            & events appended to in-memory journal \\
%  v\_apply           & journal volatily applied to in-memory metadata store \\
%  save              & journal saved to client's disk \\
%  persist           & journal saved in object store \\
%  replay            & metadata servers materialize namespace \\
%  RPCs              & round trip remote procedure calls \\
%  stream            & metadata server streams journal into RADOS \\
%\end{tabular}

\subsection{Setting Policies with Cudele}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete ({\it i.e.} the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that adminstrators can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``eventual" merges updates at some time in the
future ({\it e.g.}, when the system has time, when number of updates reaches a
certain threshold, when the client is done writing, etc.); ; and updates in
``global" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability; DeltaFS uses ``invisible" consistency
and ``local" durability and BatchFS uses ``eventual" consistency and ``local"
durability. These are just a few of the HPC examples.

% how its done
To compose the mechanisms administrators inject which steps (described in
Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use in parallel
using a domain specific language. For example, to get the semantics of BatchFS,
the administrator would inject the following pipeline:\\

\noindent \texttt{create+save+volatile apply}\\

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since ``global" durability is stronger and has
more overhead than ``local" durability.  Formulaically, valid steps
include:\\

\noindent \texttt{<create|RPCs> [+v\_apply][+save][+persist][+stream]}

\subsection{Cudele Namespace API:\\Per-Subtree Policies}
\label{sec:cudele-namespace-api-per-subtree-policies}

@@

would decouple the path \texttt{msevilla/mydir} and would apply the policies in
\texttt{policies.txt}. The policies file supports the following values:

@@

  \item \texttt{interfere\_callback}: how to handle a request from another
  client targeted at the now decoupled subtree (default \texttt{block})

@@

Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. Table~\ref{table:spectrum} shows how one would use the Cudele
API to implement policies from related work.

%

For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources for updates
that may get overwritten. If the application does not mind losing updates, for
example it wants approximations for results that take too long to compute, it
can select \texttt{overwrite}. In this case, metadata will be written and the
computation from the decoupled namespace will take priority at merge time because the results
are more accurate.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create"
    "durability": "local_persist"
}
\end{minted}
\caption{Implementing DeltaFS with Cudele.}
\label{src:deltafs}
\end{listing}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create+volatile_apply"
    "durability": "local_persist"
}
\end{minted}
\caption{Implementing BatchFS with Cudele.}
\label{src:batchfs}
\end{listing}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "-"
    "interfere_policy": "-"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\caption{Existing CephFS on Cudele.}
\label{src:batchfs}
\end{listing}

\subsection{Programmability Approach}

% what is programmability
A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach and re-uses some of the building blocks from the Malacology
programmable storage system~\ref{sevilla:eurosys17} to great success and
requires only:

  \item 354 lines of library code

  \item 219 lines of non-destructive metadata server code, which is not used
  unless it is turned on

  \item 4 lines of destructive client/server code to check whether a namespace
  is decoupled

% how it works in CephFS, how I used it
\subsubsection{Metadata Store}

\subsubsection{Journal}

\subsubsection{Journal Tool}
The journal tool is used for disaster recovery and lets administrators view and
modify the journal of metadata updates; it can read the journal, erase events
from the journal, and apply the updates in the journal to the metadata store.
To apply journal updates to the metadata store, the journal tool reads the
journal segments from object store objects and applies the update to the
metadata store (which are also stored as object store objects).

% Journal import
The journal tool imports journals from binary files stored on disk.  First the
header of the dump is sanity checked and written into RADOS to the ``header"
object.  The ``header" object has metadata about the journal as well as the
locations of all the journal pointers (e.g., where the tail of the journal is,
where we are currently trimming, etc.).  Then the journal events are cleaned
(erasing trailing events that are not part of the header) and written as
objects into RADOS.  Note that while the journal is in RADOS, the metadata
servers do do not have the namespace reconstructed in memory so the metadata
cluster will not service requests relating to the journal of imported events.
To construct the namespace in the collective memory of the metadata servers we
need to first construct the namespace in RADOS. The journal tool can explicitly
do this by  applying the journal to the metadata store in RAODS. This will pull
the objects containing journal segments and replay them on the metadata store.
Finally, we delete the journal in RADOS and restart the metadata servers so
they rebuild their caches.

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them.

% Journal export
When exporting a journal of events, the journal tool first scans the journal to
check for corruption. Then it recovers the journal by reading the ``header"
object out of RADOS.  After reading the header, the journal tool can pull
journal segments from RADOS because it knows how many objects to pull and how
far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

\subsubsection{Inode Cache}

\subsubsection{Large Inodes}

% file type
To assign consistency and durability to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{sevilla:eurosys17} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.

% lead into implementation

% // limit consistency
% if metadata write (create) -- try to reduce error bars for many creates and runtime of one touch)
%   if !RPCs
%     if overwrite
%       do nothing and let the RPC go
%     else
%       return -EBUSY
%
% // limit durability
% if !stream
%   return immediately on mdlog

>>>>

\section{Methodology: Global Namespace with Per-Subtree Consistency/Durability}
\label{sec:methodology-decoupled-namespaces}

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications decouple the namespace, write updates to a local journal,
and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
\end{figure}

In this section we describe CudeleFS, our file system that lets administrators
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. CudeleFS exposes these
mechanisms and the administrator composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

@@

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Of the 6 mechanisms in
Figure~\ref{fig:decouple} only 4 had to implemented and just 1 required changes to
the underlying storage system itself.

\subsubsection{No Changes} ``RPCs" and ``Stream" are part of the default CephFS
implementation; they are used to get strong consistency and global durability,
respectively.  Using existing configuration settings in Ceph we can turn
``Stream" on and off.  If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run.

\subsubsection{Library} for ``Create", ``Nonvolatile Apply", ``Local/Global
Persist", CudeleFS provides a library for clients to link into and all
operations are performed by the client.  Decoupled clients use the ``Create"
mechanism to append metadata updates to a local, in-memory journal.  For
``Local Persist", clients write serialized log events to a file on local disk
and for ``Global Persist", clients push the journal into the objects store. The
overheads for both ``Local Persist" and ``Global Persist" is the write
bandwidth of the local disk and object store, respectively.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  These implementations required no changes to CephFS
because the metadata servers know how to read the events the library is writing
into the object store.  By re-using the journal subsystem to implement the
namespace decoupling, Cudele leverages the write/read optimized data
structures, the formats for persisting events (similar to TableFS's
SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
onto the internal namespace data structures.

% - how it makes no gaurantees
\subsubsection{Changes to Metadata Server} the ``Volatile Apply" mechanism
takes an in-memory journal on the client and applies the updates directly to
the in-memory namespace maintained by the metadata servers. We say volatile
because -- in exchange for peak performance -- Cudele makes no consistency or
durability guarantees while ``Volatile Apply" is executing.  If a concurrent
update from a client occurs there is no rule for resolving conflicts and if the
client or metadata server crashes there may be no way to recover.

% difference between apply and volatile apply
In contrast, ``Nonvolatile Apply" uses the the object store to apply the
journal of updates from the client onto the metadata servers' metadata store.
``Nonvolatile Apply" is safer but has a performance overhead because objects in
the metadata store need to be read from and written back to the object store.

%The metadata store and journal
%are different ways of representing the namespace.  Cudele presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Create" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Create            & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{Cudele's mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

\subsection{Defining Policies in Cudele}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. 
\label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that adminstrators can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``weak" merges updates at some time in the
future ({\it e.g.}, when the system has time, when number of updates reaches a
certain threshold, when the client is done writing, etc.); and updates in
``strong" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability; DeltaFS uses ``invisible" consistency
and ``local" durability and BatchFS uses ``weak" consistency and ``local"
durability.  To compose the mechanisms administrators inject which mechanisms
(described in Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use
in parallel using a domain specific language.

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both mechanisms do the same
thing or \texttt{stream+save} since ``global" durability is stronger and has
more overhead than ``local" durability.

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed globally consistent until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that chanages a subtree from ``local
persist" to ``global persist", CudeleFS no guarantee until the mechanisms are
complete.

\subsection{Cudele Namespace API}
\label{sec:cudele-namespace-api}

@@

would decouple the path \texttt{msevilla/mydir} and would apply the policies in
\texttt{policies.yml}. The policies file supports the following values:

@@

  \item \texttt{interfere\_callback}: how to handle a request from another
  client targeted at the now decoupled subtree (default \texttt{overwrite})

@@

% description
For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources
for updates that may get overwritten. If the application does not mind losing
updates, for example it wants approximations for results that take too long to
compute, it can select \texttt{overwrite}. In this case, metadata will be
written and the computation from the decoupled namespace will take priority at
merge time because the results are more accurate.

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
would use the configuration from Listing~\ref{src:deltafs}. BatchFS's merging
back into the global namespace can be achieved with the configuration in
Listing~\ref{src:batchfs}.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create"
    "durability": "local_persist"
}
\end{minted}
\caption{Implementing DeltaFS with CudeleFS.}
\label{src:deltafs}
\end{listing}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create+volatile_apply"
    "durability": "local_persist"
}
\end{minted}
\caption{Implementing BatchFS with CudeleFS.}
\label{src:batchfs}
\end{listing}

%\begin{listing}
%\begin{minted}[frame=single,
%               framesep=3mm,
%               xleftmargin=21pt,
%               tabsize=4]{js}
%{     
%    "allocated_inodes": "-"
%    "interfere_policy": "-"
%    "consistency": "RPCs"
%    "durability": "stream"
%}
%\end{minted}
%\caption{Existing CephFS on Cudele.}
%\label{src:batchfs}
%\end{listing}

### 2017-04-01 paper/sections/5_implementation.tex
paper: integrate noah section 3 feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% Outline of the section
Each section below corresponds to the labeled arrows in
Figure~\ref{fig:decouple}. This implementation decouples policies from
mechanisms allowing applications to choose the consistency and durability
semantics they need.

% Programmability
Of the 6 mechanisms in Figure~\ref{fig:decouple} 4 had to implemented and only
1 required changes to the underlying storage system itself.  ``RPCs" and
``Stream" can be achieved with configuration settings 
.  ``Global Persist", ``Local Persist",
``Nonvolatile Apply", and ``Create" are implemented as a library and they do
not require modifications to the storage system. ``Volatile Apply" requires
changes to the metadata server to inject updates back into the global
namespace.

\subsection{No Changes: RPCs, Stream}
\label{sec:no-changes-rpcs-stream}

RPCs is the default behavior of CephFS.  Using existing configuration settings
({\it e.g.}, metadata cache size, logging on/off, etc.) in Ceph we can turn
``Stream" on and off.  If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run.

@@

\subsection{Library: Create, Local/Global Persist, Nonvolatile Apply}
\label{sec:library-create-local-global-persist}
%Reads/Writes Journal Events}

% How it works
For these mechanisms, Cudele provides a library for clients to link into.  For
``Create", clients write to an in-memory journal and updates are merged into
the global namespace by replaying them onto the metadata store either in the
metadata servers or the object store.  For ``Local Persist", clients write
serialized log events to a file on local disk and for ``Global Persist",
clients push the journal into the objects store. The overheads for both ``Local
Persist" and ``Global Persist" is the write bandwidth of the local disk and
object store, respectively. Finally, for ``Nonvolatile Apply" the journal
updates are replayed onto the metadata store in the object store and the
metadata servers are restarted.

% Level of intrusion
These implementations required no changes to Ceph because the metadata servers
know how to read the events the library is writing into the object store.  By
re-using the journal subsystem to implement the namespace decoupling, Cudele
leverages the write/read optimized data structures, the formats for persisting
events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
functions for replaying events onto the internal namespace data structures.

% Why we re-use stuff
%Step 3 is the most complicated and requires understanding how the snapshot is
%materialized in memory. 
%
%\subsubsection{Operating on Snapshots} 
%
%Our first implementation attempted to re-create journal events using the same
%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
%to instantiate a Ceph inode and directory entry for the current file/dir and
%its parent.  This is too hard because there are too many moving parts in the
%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
%traverse up and down namespace, etc.). So when I tried add dentries and inodes
%it was trying to traverse up/down and it would almost always segfault when it
%was looking for something. These metablobs are supposed to be self container --
%the problem is I do not know what is supposed to go *inside* them. 
%
%Our second idea was to copy the metadata blog and change just what we needed.
%For example, we would save a binary dump of a generic \texttt{mkdir} event on
%disk. When the application makes a directory, this dump would be loaded and the
%fields would be changed before being written back to disk. Rather than
%traversing up and down a namespace in memory of a metadata server, we should
%traverse up and down the namespace *inside* the metadata blob. This
%implementation requires disk IO and editing the log event is non-trivial for
%two reasons:
%
%\begin{itemize}
%
%  \item methods do not edit events; they just write them
%
%  \item the metadata that the event touches (e.g., the metablob) is unorganized
%  on disk for performance -- it is trade-off for writing data faster serially and
%  reconstructing information slowly since failure is not considered the norm
%
%\end{itemize}
%
%Faced with these challenges we landed on our final implementation: load the
%snapshot into the the data structures used to examine and replay journals, edit
%those data structures, and write them out to disk as binary.

\subsection{Storage System Changes: Volatile Apply}

% - how it makes no gaurantees
The ``Volatile Apply" mechanism takes an in-memory journal on the client and
applies the updates directly to the in-memory namespace maintained by the
metadata servers. We say volatile because -- in exchange for peak performance
-- Cudele makes no consistency or durability guarantees while the mechanism is
executing.  If a concurrent update from a client occurs there is no rule for
resolving conflicts and if the client or metadata server crashes there may be
no way to recover.

% difference between apply and volatile apply
In contrast, ``Nonvolatile Apply" uses the the object store to merge the journal of updates
from the client to the metadata server. ``Nonvolatile Apply" is safer but
has a performance overhead because objects in the metadata store need to be
read from and written back to the object store.  ``Volatile Apply" was already implemented
by the journal tool: it reads the journal from the object store, adds events to
the journal, and writes the metadata updates out to the metadata store in the
object store.

>>>>

A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach and re-uses some of the building blocks from the Malacology
programmable storage system~\cite{sevilla:eurosys17} to great success and
requires only:

\begin{itemize}

  \item 354 lines of library code

  \item 219 lines of non-destructive metadata server code, which is not used
  unless it is turned on

  \item 4 lines of destructive client/server code to check whether a namespace
  is decoupled

\end{itemize}

% how it works in CephFS, how I used it
\subsection{Metadata Store}
\label{sec:metadata-store}

% how it works in CephFS
In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse.\\

% how i used it
\noindent\textbf{Cudele}: uses the metadata store format to write objects
formatted like the journal into the object store. By writing with the same
format, the metadata servers can read and use the recovery code to materialize
the updates from a client's decoupled namespace ({\it i.e.} merge).

\subsection{Journal}
\label{sec:journal}

The journal is the second way that CephFS represents the file system namespace;
it is a log of metadata updates that can materialize the namespace when the
updates are replayed onto the metadata store. The journal is a ``pile system";
writes are fast but reads are slow because state must be reconstructed.
Specifically, reads are slow because there is more state to read, it is
unorganized, and many of the updates may be redundant.\\

\noindent\textbf{Cudele}: uses the journal subsystems within the metadata
server to replay updates onto the in-memory metadata store.  When the clients
are ready to merge their updates back into the global namespace, they pass a
binary file of  metadata updates to the metadata server. The metadata server
uses its recovery code to ``replay" the updates.

\subsection{Journal Tool}
\label{sec:journal-tool}

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.\\

\noindent\textbf{Cudele}: using the journal tool, clients can materialize the
journal in memory and append events to their own copy. When they are done they
export their journal to a binary file. Clients also use the journal tool to
save their updates to disk or to the object store.

% Journal import
%The journal tool imports journals from binary files stored on disk.  First the
%header of the dump is sanity checked and written into RADOS to the ``header"
%object.  The ``header" object has metadata about the journal as well as the
%locations of all the journal pointers (e.g., where the tail of the journal is,
%where we are currently trimming, etc.).  Then the journal events are cleaned
%(erasing trailing events that are not part of the header) and written as
%objects into RADOS.  Note that while the journal is in RADOS, the metadata
%servers do do not have the namespace reconstructed in memory so the metadata
%cluster will not service requests relating to the journal of imported events.
%To construct the namespace in the collective memory of the metadata servers we
%need to first construct the namespace in RADOS. The journal tool can explicitly
%do this by  applying the journal to the metadata store in RAODS. This will pull
%the objects containing journal segments and replay them on the metadata store.
%Finally, we delete the journal in RADOS and restart the metadata servers so
%they rebuild their caches.

% Journal export
%First the
%journal is scanned for the header and then journal is recovered. To recover the
%journal the ``header" object is read off disk and then objects are probed in
%order and starting from the write position saved in the header. Probing will
%update the write position if it finds objects with data in them.

% Journal export
%When exporting a journal of events, the journal tool first scans the journal to
%check for corruption. Then it recovers the journal by reading the ``header"
%object out of RADOS.  After reading the header, the journal tool can pull
%journal segments from RADOS because it knows how many objects to pull and how
%far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

\subsection{Inode Cache}
\label{sec:inode-cache}

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing operations like \texttt{lookup()}.  For example, if a client or
metadata server is not caching the directory inode, all creates within that
directory will result in a lookup and a create request. If the directory inode
is cached then only the create needs to be sent.  The size of the inode cache
is configurable so as not to saturate the memory on the metadata server --
inodes in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.

\subsection{Large Inodes}
\label{sec:large-inodes}

In CephFS, inodes store policies for policies, like how the file is striped
across the object store or for managing subtrees for load balancing. These
policies adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other.  \\

\noindent\textbf{Cudele}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

@@

%Reads/Writes Journal Events}

% How it works
% Why we re-use stuff
%Step 3 is the most complicated and requires understanding how the snapshot is
%materialized in memory. 
%
%\subsection{Operating on Snapshots} 
%
%Our first implementation attempted to re-create journal events using the same
%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
%to instantiate a Ceph inode and directory entry for the current file/dir and
%its parent.  This is too hard because there are too many moving parts in the
%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
%traverse up and down namespace, etc.). So when I tried add dentries and inodes
%it was trying to traverse up/down and it would almost always segfault when it
%was looking for something. These metablobs are supposed to be self container --
%the problem is I do not know what is supposed to go *inside* them. 
%
%Our second idea was to copy the metadata blog and change just what we needed.
%For example, we would save a binary dump of a generic \texttt{mkdir} event on
%disk. When the application makes a directory, this dump would be loaded and the
%fields would be changed before being written back to disk. Rather than
%traversing up and down a namespace in memory of a metadata server, we should
%traverse up and down the namespace *inside* the metadata blob. This
%implementation requires disk IO and editing the log event is non-trivial for
%two reasons:
%
%\begin{itemize}
%
%  \item methods do not edit events; they just write them
%
%  \item the metadata that the event touches (e.g., the metablob) is unorganized
%  on disk for performance -- it is trade-off for writing data faster serially and
%  reconstructing information slowly since failure is not considered the norm
%
%\end{itemize}
%
%Faced with these challenges we landed on our final implementation: load the
%snapshot into the the data structures used to examine and replay journals, edit
%those data structures, and write them out to disk as binary.
